---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

-----------------------------------------------------------------------------------
Title: "IST687 – Text Mining HW"
Name: Sathish Kumar Rajendiran
Week: 8
Date: 05/27/2020
-----------------------------------------------------------------------------------
Exercise: 
        Text Mining HW

# Install necessary packages

```{r}
install.packages( pkgs=c("gdata","RCurl","ggplot2","ggcorrplot","reshape2","ggeasy","viridis","viridisLite"),repos = "http://cran.us.r-project.org")

library(gdata)
library(RCurl)
library(ggplot2)
library(ggcorrplot)
library(reshape2)
library(ggeasy)
library(viridis)
library(viridisLite)

```

#printDataInfo function

```{r}
# function printDataInfo
  printDataInfo <- function(myData)
    {
      strinfo <- str(myData)
      cat("str:",strinfo,"\n")
      
      colnamesinfo <- colnames(myData)
      cat("colnames:",colnamesinfo,"\n")
      
      diminfo <- dim(myData)
      cat("dim:",diminfo,"\n")
      
      nrowinfo <- nrow(myData)
      cat("nrow:",nrowinfo,"\n")
      
      nrowsinfo <- myData[1:5,]
      return(nrowsinfo)
     }
```

#1. Read in data from the following URL:

<!-- http://college.cengage.com/mathematics/brase/understandable_statistics/7e/students/datasets/mlr/excel/mlr01.xls -->
<!-- This URL will enable you to download the dataset into excel. -->

<!-- The more general web site can be found at: -->
<!--     http://college.cengage.com/mathematics/brase/understandable_statistics/7e/students/datasets/mlr/frames/frame.html -->
    
<!-- Thunder Basin Antelope Study -->
<!-- Thunder Basin National Grassland, 2250 East Richards Street, Douglas, WY -->

<!-- The data (X1, X2, X3, X4) are for each year. -->
<!-- X1 = spring fawn count/100 -->
<!-- X2 = size of adult antelope population/100 -->
<!-- X3 = annual precipitation (inches) -->
<!-- X4 = winter severity index (1=mild,5=severe) -->


<!--     
     Antelopes are considered the sister group to deer within the infraorder Pecora.A group of antelope is called a herd.Unlike deer antlers, which are shed and grown annually, antelope horns grow continuously. 
     "Fawn" is used to describe any deer under one year of age. Most fawns are born in late spring and have lost all their spots by winter 

     Precipitation is rain, snow, sleet, or hail — any kind of weather condition where something's falling from the sky.Precipitation has to do with things falling down, and not just from the sky. It's also what happens in chemical reactions when a solid settles to the bottom of a solution.

     Winter Severity Index: 
          Winter Storm Severity Index (WSSI) is to provide National Weather Service (NWS) partners and the general public with an indication of the level of winter   precipitation (snow and ice) severity and its potential related societal impacts. The WSSI does not depict official warnings, and should always be used in context with official NWS forecasts and warnings. 
          1 -- Mild or small accumulation of snow or ice. Minimal impact | Light green 
          2 -- Minor distruptions| Advisory level | Green 
          3 -- Moderate | Warning level | Yellow 
          4 -- Major | Significant Impacts | Orange 
          5 -- Extreme | historic - Severe impacts | Red 

      for more,  https://www.weather.gov/gjt/WSSI_Tutorial
  -->
   
1. Read in data from the following URL:If you view this in a spreadsheet, you will find that four columns of a small dataset. The first column shows the number of fawn in a given spring (fawn are baby Antelope). The second column shows the population of adult antelope, the third shows the annual precipitation that year, and finally, the last column shows how bad the winter was during that year.
2. You have the option of saving the file save this file to your computer and read it into R, or reading the data directly from the web into a data frame.
3. You should inspect the data using the str() command to make sure that all of the cases have been read in (n=8 years of observations) and that there are four variables.

```{r}

# filepath

    filepath <- "http://college.cengage.com/mathematics/brase/understandable_statistics/7e/students/datasets/mlr/excel/mlr01.xls"

     # function readFiles
     readFiles <- function(fpath)
        {
        dftemp <- read.xls(filepath)
        return(dftemp)
      }
    
    
    #Import the data into temporary datasets
    
    mlrdf <- data.frame(readFiles(filepath),stringsAsFactors = FALSE)
    mlrdf[1:3]
    
    
    colnames(mlrdf) <- c("fawn","antelope","precipitation","wsi")
    #Review the dataframe
    printDataInfo(mlrdf)
    
```

4. Create bivariate plots of number of baby fawns versus adult antelope population, the precipitation that year, and the severity of the winter. Your code should produce three separate plots. Make sure the Y-axis and X-axis are labeled. Keeping in mind that the number of fawns is the outcome (or dependent) variable, which axis should it go on in your plots?

```{r}
#Baby Fawns versus Adult Antelope Population

  theme <-theme(plot.title = element_text(hjust = 0.5),axis.title = element_text())
  g1 <- ggplot(mlrdf,aes(x=antelope,y=fawn))+geom_point(aes(),size = 3,color="#FF99FF") + geom_smooth(method = "lm",color="black")
  g1 + labs(x = "adult antelope population", y = "baby fawns", 
            title = "Baby Fawns versus Adult Antelope Population", 
            caption = "Data: http://college.cengage.com/ + mlr01.xls ") + theme
```



```{r}
#Baby Fawns versus Precipitation

  g2 <- ggplot(mlrdf,aes(x=precipitation,y=fawn))+geom_point(aes(),size = 3,color="#4393C3") + geom_smooth(method = "lm",color="#F4A582")
  g2 + labs(x = "precipitation", y = "baby fawns", 
            title = "Baby Fawns versus Precipitation", 
            caption = "Data: http://college.cengage.com/ + mlr01.xls ") + theme 
  
```


```{r}
 #Baby Fawns versus severity of the winter

  g3 <- ggplot(mlrdf,aes(x=wsi,y=fawn))+geom_point(aes(),size = 3,color="#B2182B") + geom_smooth(method = "lm",color="#009999")
  g3 + labs(x = "severity of the winter", y = "baby fawns", 
            title = "Baby Fawns versus Severity of the Winter", 
            caption = "Data: http://college.cengage.com/ + mlr01.xls ") + theme 
```
#5. Next, create three regression models of increasing complexity using lm()

```{r}
    # predict the number of fawns from the severity of the winter
   
    mlr.lm1 <- lm(formula = fawn ~ wsi,data = mlrdf)
    summ <- summary(mlr.lm1)
    
    summary(mlr.lm1)
    
    paste("p-values:")
    summ$coefficients[,4]
    
    paste("adjusted r-squared:" ,summ$adj.r.squared)
    
    # str(summ)

```

```{r}
    # predict the number of fawns from the severity of the winter + precipitation
   
    mlr.lm2 <- lm(formula = fawn ~ wsi+precipitation,data = mlrdf)
    summ <- summary(mlr.lm2)
    
    summary(mlr.lm2)
    
    paste("p-values:")
    summ$coefficients[,4]
    
    paste("adjusted r-squared:" ,summ$adj.r.squared)
    
    # str(summ)
```
```{r}
   # predict the number of fawns from the severity of the winter + antelope
   
    mlr.lm3 <- lm(formula = fawn ~ wsi+antelope,data = mlrdf)
    summ <- summary(mlr.lm3)
    
    summary(mlr.lm3)
    
    paste("p-values:")
    summ$coefficients[,4]
    
    paste("adjusted r-squared:" ,summ$adj.r.squared)
    
    # str(summ)
```



```{r}

     # predict the number of fawns from the severity of the winter + precipitation + antelope
   
    mlr.lm4 <- lm(formula = fawn ~ wsi+precipitation+antelope,data = mlrdf)
    summ <- summary(mlr.lm4)
    
    summary(mlr.lm4)
    
    paste("p-values:")
    summ$coefficients[,4]
    
    paste("adjusted r-squared:" ,summ$adj.r.squared)
    
    
```

```{r}

   
# Conclusion: predict the number of fawns from the severity of the winter + precipitation has better Adjusted R square (86%) other combinations.
   
```

