---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

-----------------------------------------------------------------------------------
Title: "IST687 – Cleaning/munging Dataframes"
Name: Sathish Kumar Rajendiran
Week: 3
Date: 04/20/2020
-----------------------------------------------------------------------------------
Exercise: 
        Data munging on state populations (within the United States).


Step 1: Create a function (named readStates) to read a CSV file into R

      * Note that you are to read a URL, not a file local to your computer.
      * The file is a dataset on state populations (within the United States).
      * The URL is: http://www2.census.gov/programs-surveys/popest/tables/2010-2011/state/totals/nst-est2011-01.csv

```{r}

#   1. Note that you are to read a URL, not a file local to your computer.
#   2. The file is a dataset on state populations (within the United States).
      
      #file path
      fpath <- "http://www2.census.gov/programs-surveys/popest/tables/2010-2011/state/totals/nst-est2011-01.csv"
     
      # function readStates
      readStates <- function(fpath)
        {
        dftemp <- read.csv(url(fpath),skip=3)
        #head(dfStates,10)
        #colnames(dfStates)
        return(dftemp)
        }
      
      dftemp <- readStates(fpath) 
      head(dftemp)
  
```

Step 2: Clean the dataframe

```{r}

#   3. Note the issues that need to be fixed (removing columns, removing rows, changing column names).
      
    head(dftemp)
    tail(dftemp)
    summary(dftemp)
    
#   4. Within your function, make sure there are 51 rows (one per state + the district of following names 
#      (stateName,base2010,base2011,Jul2010,Jul2011)

    # remove empty column
    dftemp <- dftemp[,c(1,2,3,4,5)]
    #head(dfStates)

    colnames(dftemp)<- c('stateName', 'base2010', 'base2011','Jul2010', 'Jul2011')
    #summary(dfStates)
    
    # Keep only rows corresponding to the states & remove empty rows
    dftemp <- dftemp[6:56,]
    #head(dfStates)

#   5. Make sure the last four columns are numbers (i.e. not strings)
    dftemp$base2010 <- as.integer(gsub(',','',dftemp$base2010))
    dftemp$base2011 <- as.integer(gsub(',','',dftemp$base2011))
    dftemp$Jul2010 <- as.integer(gsub(',','',dftemp$Jul2010))
    dftemp$Jul2011 <- as.integer(gsub(',','',dftemp$Jul2011))
    
    # remove the special character "." infront of stateName
    dftemp$stateName <- gsub('^[.]{1}','',dftemp$stateName)
    
    # reset Row Index
    row.names(dftemp) <- NULL
    
    summary(dftemp)
    head(dftemp)
    
```

Step 3: Store and Explore the dataset
        Putting all together into one function
```{r}

#     6. Store the dataset into a dataframe, called dfStates

      # file path
      fpath <- "http://www2.census.gov/programs-surveys/popest/tables/2010-2011/state/totals/nst-est2011-01.csv"

      # function readStates
      readStates <- function(fpath)
        {
        dftemp <- read.csv(url(fpath),skip=3)
        dftemp <- dftemp[,c(1,2,3,4,5)]
        colnames(dftemp)<- c('stateName', 'base2010', 'base2011','Jul2010', 'Jul2011')
        dftemp <- dftemp[6:56,]
        dftemp$base2010 <- as.integer(gsub(',','',dftemp$base2010))
        dftemp$base2011 <- as.integer(gsub(',','',dftemp$base2011))
        dftemp$Jul2010 <- as.integer(gsub(',','',dftemp$Jul2010))
        dftemp$Jul2011 <- as.integer(gsub(',','',dftemp$Jul2011))
        dftemp$stateName <- gsub('[.]','',dftemp$stateName)
        
        row.names(dftemp) <- NULL
        return(dftemp)
        }
      
      dfStates <- readStates(fpath) 
      head(dfStates)
      
#     7. Test your dataframe by calculating the mean for the July2011 data, by doing:
      
      Jul2011Mean <- mean(dfStates$Jul2011)
      sprintf("%.0f is the Jul2011 mean",Jul2011Mean)
      
```

Step 4: Find the state with the Highest Population

```{r}

#     8. Based on the July2011 data, what is the population of the state with the highest population? What is the name of that state?
      
      maxPopulation <- max(dfStates$Jul2011)
      sprintf("%.0f is the highest population in July 2011",maxPopulation)
      
      maxPopState_index <- which.max(dfStates$Jul2011)
      maxPopState <- dfStates[maxPopState_index,1]
      print(paste("State with highest population in July 2011 is:",maxPopState),max.levels = 0)

#     9. Sort the data, in increasing order, based on the July2011 data
      
      dfStates <- dfStates[order(dfStates$Jul2011),]
      head(dfStates)
      tail(dfStates)

```

Step 5: Explore the distribution of the states

```{r}
#     10. Write a function that takes two parameters. The first is a vector and the second is a number.
#     11. The function will return the percentage of the elements within the vector that is less than the same (i.e. the cumulative distribution below the value provided).
#     12. For example, if the vector had 5 elements (1,2,3,4,5), with 2 being the number passed into the function, the function would return 0.2 (since 20% of the numbers were below 2)
      v <- c(1,2,3,4,5)

# function readStates
      cumDist <- function(v,n)
        {
        #distValue <- mean(v<n)  #--option using mean
        distValue <- prop.table(table(v < n))[2]  #-- option using proportion
        #distValue <- length(which(v<n))/length(v)  # option using length
        return(distValue)
        }
      
      distValue <- cumDist(v,2) 
      distValue_percent <- distValue*100
      sprintf("%0.0f Percentage of population are less than the average population in July 2011",distValue_percent)
```

13. Test the function with the vector ‘dfStates$Jul2011Num’, and the mean of dfStates$Jul2011Num’.
There are many ways to write this function (described in #10 above) – so please try to write multiple versions of this function – which do you think is best?

```{r}
#     13. Test the function with the vector ‘dfStates$Jul2011Num’, and the mean of dfStates$Jul2011Num’
        cumDist <- function(v,n)
          {
            #distValue <- prop.table(table(v < n))[2]  #-- option using proportion
            #distValue <- length(which(v<n))/length(v)  # option using length
            distValue <- mean(v<n)  #--option using mean
            return(distValue)
          }
        meanJul2011 <- mean(dfStates$Jul2011)
        distValue <- cumDist(dfStates$Jul2011,meanJul2011) 
        distValue_percent <- distValue*100
        sprintf("%.0f Percentage of population are less than the average population in July 2011",distValue_percent)
       
         #67 Percentage of population are less than the average population in July 2011" This option looks simpler
```

