---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
-----------------------------------------------------------------------------------
Title: "IST687 – Final Project | 2016 Olympics in Rio de Janeiro"
Name: Sathish Kumar Rajendiran
Group Number: 3
Members: Deanna Zarrillo, Jeff Thomson, Sathish Kumar Rajendiran
Date: 05/12/2020
-----------------------------------------------------------------------------------
Exercise: 
        2016 Olympics, Rio de Janeiro, Brazil

Countries:
Number of Players:
Male:
Female:
Medal Category: Gold, Silver, Bronze
Games:
Venues Played:

￼
1. Problem formulation
2. Data cleanup
3. Pre-processing
4. Train-test Split
5. Model Building
6. Accuracy & Validation
7. Prediction
8. Confusion Matrix

Descriptive Analysis
Predictive Analysis
Prescrptive Analysis

# Install necessary packages

```{r}
# import libraries

install.packages(pkgs=c("ggplot2","reshape2","ggeasy","dplyr","data.table","RSQLite","sqldf","reshape"),repos = "http://cran.us.r-project.org")

library(ggplot2)
library(ggeasy)
library(reshape2)
library(data.table)
library(stats)
library(dplyr)
library(RSQLite)
library(sqldf)
library(reshape) 
library(tidyverse)
library(rvest)
library(magrittr)
library(ggmap)
library(stringr)
library(RSQLite)
library(viridis)
library(hrbrthemes)

```


#1. Import raw data into dataframes
    1. Download files
    2. Create data frames
    3. Review meta data
 
<!-- https://www.kaggle.com/rio2016/olympic-games -->

```{r}

#import csv files into data frames

  athletes_fpath <- "/Users/sathishrajendiran/Documents/olympic-games/athletes.csv"
  countries_fpath <- "/Users/sathishrajendiran/Documents/olympic-games/countries.csv"
  events_fpath <- "/Users/sathishrajendiran/Documents/olympic-games/events.csv"

  # function readFiles
  readFiles <- function(fpath)
    {
    dftemp <- data.frame(read.csv(fpath),stringsAsFactors=FALSE)
    return(dftemp)
    }
  
  athletes <- readFiles(athletes_fpath) 
  countries <- readFiles(countries_fpath) 
  events <- readFiles(events_fpath) 

  
# function printDataInfo
  printDataInfo <- function(myData)
    {
      strinfo <- str(myData)
      cat("str:",strinfo,"\n")
      
      colnamesinfo <- colnames(myData)
      cat("colnames:",colnamesinfo,"\n")
      
      diminfo <- dim(myData)
      cat("dim:",diminfo,"\n")
      
      nrowinfo <- nrow(myData)
      cat("nrow:",nrowinfo,"\n")
      
      nrowsinfo <- myData[1:3,]
      return(nrowsinfo)
     }
  
# printDataInfo
  printDataInfo(athletes)  #dim: 11538 11
  printDataInfo(countries) #dim: 201 4
  printDataInfo(events) #dim: 306 6 
  
  
 
```


#2. Data Cleaning Process

    1. Remove NA Characters
    2. Change Data type 
    3. Remove unwanted columns
    4. Add columns (Copy + calculated)
    5. Rename Columns
    6. Re-arrange Columns
    7. ?


 2.1 Athletes Dataframe


```{r}
# 2. Data Cleaning Process | 2.1 Athletes Dataframe

#   Preview the data
    
    printDataInfo(athletes)
    
    
      # 2.1.1   [ No need now]
              #  Data format of dob column is as factor
              #  Lets change it to yyyy-mm-dd format
          
              # athletes$dob = as.Date(athletes$dob,"%m/%d/%y")

# 2.1.2
    #  Check if the NAs in athletes and replace them  

    clnames <- colnames(athletes)[colSums(is.na(athletes)) > 0]
    clnames

    
    #  create subset of dataframe rows having NAs
    na_athletes <- athletes[rowSums(is.na(athletes)) > 0,]
    na_athletes # 680- rows
    
    # nasql_athletes <- sqldf("select name,nationality,dob,height,weight from na_athletes order by name")
    # nasql_athletes <- sqldf("select name,nationality,dob,height,weight from na_athletes where dob is null order by name")
    # nasql_athletes
   
    
# 2.1.3    
    # Replace dob, height & weight column's NA values
    
    # athletes[is.na(athletes$dob),]
    # athletes[which(athletes$dob==""),]
    athletes[athletes$nationality=="RUS" & athletes$name=="Pavel Sozykin",]$dob <-"12/25/1987"  # Update DOB of "Pavel Sozykin" born 25 December 1987
    
    #[ No need for now]
        # athletes[athletes$nationality=="HUN" & athletes$name=="Adam Decker",]$dob <-"1984-02-29"  # Update DOB of "Adam DECKER" born 29 Feb 1984
        # athletes[athletes$nationality=="UKR" & athletes$name=="Artem Morozov",]$dob <-"1980-02-29"  # Update DOB of "Artem Morozov" born 29 Feb 1980
        # athletes[athletes$nationality=="GBR" & athletes$name=="Hannah Mills",]$dob <-"1988-02-29"  # Update DOB of "Hannah Mills" born 29 February 1988
        # athletes[athletes$nationality=="SRB" & athletes$name=="Jovana Crnogorac",]$dob <-"1992-02-29"  # Update DOB of "Jovana Crnogorac" born 29 February 1992
        # athletes[athletes$nationality=="IND" & athletes$name=="Prakash Nanjappa",]$dob <-"1976-02-29"   # Update DOB of "Prakash Nanjappa" born 29 February 1976
        # athletes[athletes$nationality=="IND" & athletes$name=="Prakash Nanjappa",]
        # athletes[athletes$nationality=="RUS" & athletes$name=="Pavel Sozykin" ,]   # validate other RUS athletes
        
# 2.1.4
   
    # Check if the NAs in column "weight" and replace them by the mean value of this column
    athletes$weight[is.na(athletes$weight)] <- mean(athletes$weight, na.rm=TRUE)

# 2.1.5
  
    # Check if the NAs in column "height" and replace them by the mean value of this column
    athletes$height[is.na(athletes$height)] <- mean(athletes$height, na.rm=TRUE)   
    
    #  Check if the NAs in athletes and replace them   | Verify if NAs has been replaced 
    # na_clnames <- colnames(athletes)[colSums(is.na(athletes)) > 0]
    # na_clnames
    # 
 
# 2.1.5  
    
    #id Column in athletes doesnt mean much. lets remove it
    athletes <- athletes[,!grepl("id",colnames(athletes))]
    # athletes[1:3,]
 
# 2.1.6  
  
    #Rename name,nationality columns 
    colnames(athletes) <- c("fullname","country_code","sex","dob","height","weight","sport","gold","silver","bronze")
    # athletes[1:3,]

# 2.1.7 
    
    # Add columns age,tot_medals,weight_lbs,bmi_metric
    # lets add Age as a calculated column to hold age of athletes. 
    # This is by simply performing a data difference between athletes' dob and Jan 1st of 2016.
    # athletes$age <- 0
    
    athletes$age <- round(as.numeric(as.Date(as.character("01/01/2016"), format="%m/%d/%Y") - as.Date(as.character(athletes$dob), format="%m/%d/%Y"))/365,0)
    
    #  lets calculate total number of medals by summing all medal categories (gold,silver and bronze)
    athletes$tot_medals = athletes$gold+athletes$silver+athletes$bronze

    #  convert players wieght into lbs
    athletes$weight_lbs = 2.2 * athletes$weight
 
    #  calculate BMI using metric formula
    height_m2 <- ifelse(athletes$height >0, athletes$height * athletes$height,0)
    athletes$bmi_metric = round(athletes$weight/height_m2,1)
    
    # athletes[1:3,]

# 2.1.8  
    
    # re-order the columns

    ath_col_order <- c("fullname","sex","country_code","sport","dob","age","height","weight","weight_lbs","bmi_metric","gold","silver","bronze","tot_medals")
    athletes <- athletes[, ath_col_order]
    athletes[1:3,]
    
    # athletes[which(athletes$age<0),]
 
```


```{r}
#Age validation
# athletes[which(athletes$age>0),]
# athletes[which(athletes$age<0),]
# # athletes$age
# athletes[which(athletes$dob==""),]
# athletes[is.na(athletes$age),]
# 
# athletes[athletes$country_code=="RUS" & athletes$fullname=="Pavel Sozykin",]
# athletes$age <- round(as.numeric(as.Date(as.character("01/01/2016"), format="%m/%d/%Y") - as.Date(as.character(athletes$dob), format="%m/%d/%Y"))/365,0)
# athletes

    summary(athletes$age)
    head(athletes[order(athletes$age),])
    head(athletes[order(-athletes$age),])

```

#Charts definition
```{r}

# barplot(athletes$dob)
    hcolor <- c("orange")
    hfill <- c("steelblue")
    htitle <- c("values distribution")
    theme <-theme(plot.title = element_text(hjust = 0.5),axis.title = element_text())
  
   #valdiation of player's age across 
    ggAgeboxplot <- ggplot(athletes,aes(age)) +geom_boxplot(fill = "steelblue", colour = "orange",outlier.colour = "red", outlier.shape = 1,na.rm=TRUE)+ggtitle("Player's age analysis")
    
    ggAgeboxplot+theme
    


```

```{r}
#Age by gender
boxplot(age~sex,
       data = athletes, main="Player's age by gender", 
       ylab="",
       xlab = "Age",
       border="orange", 
       col="steelblue",
       freq=FALSE,
       las=1, 
       breaks=5,
       horizontal = TRUE,na.rm=TRUE)

boxplot(age~tot_medals,
       data = athletes, main="Number of medals by Age", 
       ylab="",
       xlab = "",
       border="orange", 
       col="steelblue",
       freq=FALSE,
       las=1, 
       breaks=5,
       horizontal = TRUE,na.rm=TRUE)


boxplot(tot_medals~sex,
       data = athletes, main="Number of medals by Age & Gender", 
       ylab="",
       xlab = "",
       border="orange", 
       col="steelblue",
       freq=FALSE,
       las=1, 
       breaks=5,
       horizontal = TRUE,na.rm=TRUE)


```


#2. Data Cleaning Process | 2.2 Countries Dataframe

    1. Remove NA Characters
    2. Change Data type 
    3. Remove unwanted columns
    4. Add columns (Copy + calculated)
    5. Rename Columns
    6. Re-arrange Columns
    7. ?

```{r}


#   Preview the data
    
    
    printDataInfo(countries)

# 2.2.1  
  
    #Rename code to country_code 
    colnames(countries) <- c("country","country_code","population","gdp_per_capita")
    # countries[1:3,]
    
    
# 2.2.2
   
    #  Check if the NAs in athletes and replace them  
    nacountry_clmns <- colnames(countries)[colSums(is.na(countries)) > 0]
    nacountry_clmns

    #  create subset of dataframe rows having NAs
    na_countries <- countries[rowSums(is.na(countries)) > 0,]
    na_countries # 25- rows
    
    # nasql_countries <- sqldf("select country,country_code,population,gdp_per_capita from na_countries order by country")
    # nasql_countries #25 rows

# 2.2.3
    
    # find out if any country_code from athletes dataframe  is missing in countries dataframe
    countries[1:3,]
    athletes[1:3,]
    
    x <- data.frame(countries[,1:2],stringsAsFactors=FALSE)
    y <- data.frame(athletes$country_code,stringsAsFactors=FALSE)
    colnames(y) <- c("country_code")
    z <- data.frame(unique(anti_join(y,x, by="country_code",),stringsAsFactors=FALSE))
    z

    country_code <- c("IOA","KIR","KOS","MHL","MNE","ROT","ROU","SRB","SSD","TTO","TUV")
    country <- c("Individual Olympic athletes","Kiribati","Kosovo","Marshall Islands","Montenegro","Refugee Olympic Team","Romania","Serbia","South Sudan","Trinidad and Tobago","Tuvalu")
    population <- rep(0, 11)
    gdp_per_capita <- rep(0, 11)

    missing_countries <- data.frame(country_code,country,population,gdp_per_capita,stringsAsFactors=FALSE)
    missing_countries

    countries <- rbind(countries, missing_countries)

    countries[1:3,]

    clnames3 <- colnames(countries)[colSums(is.na(countries)) > 0]
    clnames3
```


```{r }
options(scipen=999) 
 #valdiation of player's age across 


    ggPopluationboxplot <- ggplot(countries,aes(population)) +geom_boxplot(fill = "#69b3a2", colour = "steelblue",outlier.colour = "red", outlier.shape = 1,na.rm=TRUE,orientation=NA)+ggtitle("boxplot - population spread")
    
    ggPopluationboxplot+theme
    
    
  
```

#3. Join Athletes and Countries data frame

    Join Athletes and Countries data frame
    Re-arrange the columns in Athletes data frame

```{r}
#3.1
    #jojn Athletes and Countries dataframe by nationality
    
    # printDataInfo(athletes)
    # printDataInfo(countries)

   athletes <- left_join(athletes,countries, by="country_code")

#3.2
    # re-order the columns
    # colnames(athletes)
    
    col_order <- c("fullname","sex","country","country_code","sport","dob","age","height","weight","weight_lbs","bmi_metric","gold","silver","bronze","tot_medals","population","gdp_per_capita")
    athletes <- athletes[, col_order]
    # athletes[1:3,]
    
#3.3
    # review for NAs
    na_athletes <- colnames(athletes)[colSums(is.na(athletes)) > 0]
    # na_athletes
    
    #create subset of dataframe rows having NAs
    # na_athletes <- athletes[rowSums(is.na(athletes)) > 0,]
    # na_athletes #509 rows
    
    # nasql_athletes <- sqldf("select fullname,country,dob,height,weight,weight_lbs,bmi_metric,population,gdp_per_capita from na_athletes order by fullname")
    # nasql_athletes
    
    printDataInfo(athletes)
```


```{r}
# sort(unique(events$sport))

   # Values in Events and Athlets are at different levels; So, join doesnt really work between these two dataframes. 
         # events[1:3,]
    
          #jojn Athletes and events dataframe by sport & id

          # athletes <- left_join(athletes,events)
          # athletes[1:3,]
   

```


```{r}
# Plot histogram on countries data to analyze the data spread for missing data elements

  hcolor <- c("orange")
  hfill <- c("steelblue")
  htitle <- c("Histogram - Data availability")
  theme <-theme(plot.title = element_text(hjust = 0.5),axis.title = element_text())
  
  gghist <- ggplot(data=melt(athletes),mapping = aes(x= value)) 
  gghist+geom_histogram(bins = 20,color=hcolor,fill=hfill,na.rm = TRUE)+facet_wrap(~variable,scales = "free_x")+ ggtitle(htitle) + theme
    
  
  
```


```{r}
athletes[1:3,]
```

```{r}

hmap <- ggplot(athletes, aes(x=sex, y=age))
hmap <- hmap + geom_tile(aes(fill=tot_medals), color="white",na.rm=TRUE) 
hmap <- hmap + scale_fill_gradient(low= "lightblue", high="red") + ggtitle("total medals by age and sex")
hmap <- hmap + theme(
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)
hmap <- hmap + theme(
  panel.background = element_rect(fill = "black")
) 
hmap 
```


```{r}
h <- ggplot(athletes, aes(x=sex, y=age))
h <- h + geom_tile(aes(fill=bmi_metric), color="white", na.rm=TRUE)
h <- h + scale_fill_gradient(low="lightblue", high= "red")
h <- h + theme(
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)
h <- h + theme(
  panel.background = element_rect(fill = "black")
) + ggtitle("BMI by age and sex")
h
```

```{r}
p2 <- ggplot(athletes, aes(x=sex, y=bmi_metric))
p2 <- p2 + geom_tile(aes(fill=tot_medals), color="white",na.rm=TRUE) 
p2 <- p2 + scale_fill_gradient(low= "lightblue", high="red") + ggtitle("total medals by sex and BMI")
p2 <- p2 + theme(
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)
p2 <- p2 + theme(
  panel.background = element_rect(fill = "black")
) 
p2 
```


```{r}

athletes[1:3,]
m <- ggplot(athletes, aes(x=country, y=age))
m <- m + geom_tile(aes(fill=bmi_metric), na.rm=TRUE)
m <- m + scale_fill_gradient(low="lightblue", high= "red") 
m <- m + theme(axis.text.x = element_text(size=4, angle = 90))
m <- m+ theme(
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)
m <- m + theme(
  panel.background = element_rect(fill = "black")
)
m
```


```{r}
p <- ggplot(athletes, aes(x=bmi_metric, y=age))
p <- p + geom_tile(aes(fill=tot_medals), color="white",na.rm=TRUE) 
p <- p + scale_fill_gradient(low= "lightblue", high="red") + ggtitle("total medals by age and BMI")
p <- p + theme(
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
)
p <- p + theme(
  panel.background = element_rect(fill = "black")
) 
p 

```



```{r}
    
    athletes <- na.omit(athletes)  # Remove NAs from the data
    
    sqlResults <- sqldf("select  * from athletes limit 5")
    # sqlResults
    
    nbrofplayers <- sqldf("select country, sex, count(fullname) nbr_players  from athletes group by country,sex order by nbr_players desc "  )
    # nbrofplayers
    
    topcountries <- sqldf("select  country, count(fullname) nbr_players  from athletes group by country order by nbr_players desc limit 10"  )
    # topcountries
    
    
    toplevelcountries <- sqldf("select country, sum(gold) as nbr_gold, sum(silver) as nbr_silver,sum(bronze)  as nbr_bronze,sum(tot_medals) as nbr_medals from athletes group by country order by nbr_medals desc limit 10"  )
    # toplevelcountries
    
    topplayers <- sqldf("select fullname,sex, sum(gold) as nbr_gold, sum(silver) as nbr_silver,sum(bronze)  as nbr_bronze,sum(tot_medals) as nbr_medals from athletes group by fullname,sex order by nbr_medals desc"  )
    # topplayers
    
    topsports <- sqldf("select sport,count(fullname) nbr_players,sum(tot_medals) as nbr_medals from athletes group by sport order by nbr_medals desc,nbr_players desc limit 10"  )
    # topsports
    
    topmedals <- sqldf("select  country, count(fullname) nbr_players ,sum(tot_medals) as tot_medals from athletes group by country order by tot_medals desc,nbr_players desc limit 10"  )
    # topmedals
    
    topgender <- sqldf("select sport,sex, sum(tot_medals) as nbr_medals from athletes group by sport,sex order by nbr_medals desc"  )
    # topgender
    
```
```{r}

# topcountries

Countries <- topcountries$country [order (topcountries$nbr_players, decreasing = TRUE)]
# Countries

x <- subset(nbrofplayers,nbrofplayers$country %in% Countries)
       
ggplot (data = x,
            aes (nbr_players,country)) +
            geom_point (aes (color = sex), size = 5)

```


```{r}
options(scipen=999)  # turn-off scientific notation like 1e+48
# theme_set(theme_bw())  # pre-set the bw theme.


# Scatterplot
gg <- ggplot(topmedals, aes(x=tot_medals, y=nbr_players)) + 
  geom_point(aes(col=country, size=tot_medals)) + 
  geom_smooth(method="loess", se=F) + 
  xlim(c(0, 300)) + 
  ylim(c(0, 600)) + 
  labs( y="Number of Players", 
       x="Number of Medals", 
       title="Number of Players Vs Number of Medals", 
       caption = "Source: Olympic 2016")

plot(gg)
```


```{r}


topgender <- sqldf("select sport,sex, sum(tot_medals) as medals from athletes group by sport,sex order by medals desc"  )
topgender
    
str(topgender)

ggplot(topgender, aes(fill=sex, y=medals, x=sport)) + geom_bar(position="stack", stat="identity")+
    scale_color_brewer(palette = "Paired")+
    ggtitle("Sports vs Players's Gender vs Number of Goals") +
    theme(axis.text.x = element_text(face="plain", color="black", size=12, angle=90))
  
```

```{r}


 
# Create dataset
data <- data.frame(
  individual=paste( "Mister ", seq(1,60), sep=""),
  value=sample( seq(10,100), 60, replace=T)
)
 # data
 # athletes
data <- sqldf("select sport as individual, sum(tot_medals) as value from athletes  group by sport order by sport"  )

# data
    
# Set a number of 'empty bar'
empty_bar <- 10
 
# Add lines to the initial dataset
to_add <- matrix(NA, empty_bar, ncol(data))
colnames(to_add) <- colnames(data)
data <- rbind(data, to_add)
data
data$id <- seq(1, nrow(data))
 
# Get the name and the y position of each label
label_data <- data
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)
 
# Make the plot
p <- ggplot(data, aes(x=as.factor(id), y=value)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  geom_bar(stat="identity", fill=alpha("green", 0.4)) +
  ylim(-100,350) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  coord_polar(start = 0) + 
  geom_text(data=label_data, aes(x=id, y=value+5, label=individual, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=3.2, angle= label_data$angle, inherit.aes = FALSE,na.rm = TRUE ) 
 
p

```

```{r}

df1 <- sqldf("select fullname, gold, silver,bronze from athletes")  #--where fullname like 'Michael Phelps
# df1


mdata <- melt(df1, id=c("fullname"))
mdata <- mdata[order(-mdata$value,mdata$fullname),]
mdata <- mdata[which(mdata$value>=1),]
# mdata
# mdata[mdata$fullname =="Michael Phelps",]


# topcountries

Countries <- topcountries$country [order (topcountries$nbr_players, decreasing = TRUE)]
# Countries

x <- subset(nbrofplayers,nbrofplayers$country %in% Countries)
    
p1 <- sqldf("select fullname,sum(tot_medals) as medals from athletes group by fullname order by medals desc")
p1 <- p1[which(p1$medals>=3),]


players <- p1$fullname[order(p1$medals, decreasing = TRUE)] 
# players

x <- subset(mdata,mdata$fullname %in% players)
x

v <- ggplot(x)+aes(x=fullname,y = value,fill =variable)+geom_col()+ coord_flip() + scale_colour_brewer(palette = "Set1")+ theme_dark()
v

```

```{r}
# options(scipen=999) 
#  #valdiation of player's age across 
# 
# 
#     ggPopluationboxplot <- ggplot(countries,aes(population)) +geom_boxplot(fill = "#69b3a2", colour = "steelblue",outlier.colour = "red", outlier.shape = 1,na.rm=TRUE,orientation=NA,xmin = min(x), lower = quantile(x, 0.25), middle = median(x), upper = quantile(x, 0.75), xmax = max(x),stat = "identity")+ggtitle("boxplot - population spread")
#     
#     ggPopluationboxplot+theme
    
```



```{r}
# Create new table that sums the total medals by country
# Note: Some code used from this website: https://www.r-bloggers.com/how-to-make-a-global-map-in-r-step-by-step/

Top15<- athletes %>%
  group_by(country) %>%
  summarise(tot_medals = sum(tot_medals))

#Order from highest to lowest
Top15<- Top15[order(-Top15$tot_medals),]

#Only take the top 15
Top15<- Top15[1:15,]

#create map showing total medals per country
# Updates country in Top15 to Character
Top15<- Top15 %>% mutate_if(is.factor, as.character)

#Update Country Names for United States to USA and United Kingdom to UK
Top15$country <- recode(Top15$country,'United States' = 'USA','United Kingdom' = 'UK')

Top15$rank<- c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15) #add ranking order to use for colouring map by most medals

#get world map
world_map<-map_data("world")
ggplot(world_map, aes(x=long, y = lat, group=group))+
  geom_polygon(fill="lightgray", colour="black")

world.map.medals<-left_join(world_map,Top15,by=c('region'='country'))

world.map.medals<-world.map.medals %>% mutate(fill_flg = ifelse(is.na(rank),F,T))

Top15map<- ggplot() +
  geom_polygon(data = world.map.medals, aes(x = long, y = lat, group = group, fill = fill_flg,colour=rank)) +
  scale_fill_hue(l=40,c=35) +
  labs(title = '2016 Olympics, Top15 Country with Most Total Medals')
Top15map
```

```{r}
#Scatterplot of Sport, Age and Total Medals ranked by Top 10 sports with most total medals
dfAgeSportMedals<- athletes %>%
  group_by(sport) %>%
  summarise(age = mean(age),tot_medals = sum(tot_medals))
dfAgeSportMedals<- dfAgeSportMedals[order(-dfAgeSportMedals$tot_medals),]
dfAgeSportMedals<- dfAgeSportMedals[1:10,]
dfAgeSportMedals

ggplot(dfAgeSportMedals,aes(x=sport,y=age))+geom_point(aes(size=tot_medals,color=tot_medals))

```


```{r}

#Scatterplot of Sport, BMI and Total Medals ranked by Top 10 sports with most total medals
dfbmiSportMedals<- athletes %>%
  group_by(sport) %>%
  summarise(bmi = mean(bmi_metric),tot_medals = sum(tot_medals))
dfbmiSportMedals<- dfbmiSportMedals[order(-dfbmiSportMedals$tot_medals),]
dfbmiSportMedals<- dfbmiSportMedals[1:10,]
dfbmiSportMedals

#Scatterplot of Sport, Age and Total Medals
ggplot(dfbmiSportMedals,aes(x=sport,y=bmi))+geom_point(aes(size=tot_medals,color=tot_medals))

```




```{r}
#1.Regression Analysis- Age versus Total Medals

Age_totMedals<-athletes %>%
  group_by(age) %>%
  summarise(tot_medals = sum(tot_medals))

Age_totMedals<-Age_totMedals[order(-Age_totMedals$age),]

# #Only take positive
# Age_totMedals<- Age_totMedals[1:35,] # CAN DELETE ONCE INTEGRATE SATHISH's CODE THAT FIXES NEGATIVE AGES

plot(Age_totMedals$age,Age_totMedals$tot_medals)

Reg_Age<-lm(formula = tot_medals~age,data=Age_totMedals)
summary(Reg_Age)

# Assuming an alpha of 0.05, the p-value of 0.0371 indicates that age does have a correlation to total medals won

```

```{r}

#2.Regression Analysis-BMI versus Total Medals

#sathish ----
BMI_totMedals<-athletes %>%
  group_by(bmi_metric) %>%
  summarise(tot_medals = sum(tot_medals))

BMI_totMedals<-BMI_totMedals[order(-BMI_totMedals$bmi_metric),]

plot(BMI_totMedals$bmi_metric,BMI_totMedals$tot_medals)

#sathish ----

Reg_BMI<-lm(formula = tot_medals~bmi_metric,data=BMI_totMedals)
summary(Reg_BMI)

#  Assuming an alpha of 0.05, the p-value of 0^-14 indicates that BMI does have a correlation to total medals won.  
```


```{r}
#3 Run regression for Population and Total Medals

#sathish ----
Pop_totMedals<-athletes %>%
  group_by(population) %>%
  summarise(tot_medals = sum(tot_medals))

Pop_totMedals<-Pop_totMedals[order(-Pop_totMedals$population),]

plot(Pop_totMedals$population,Pop_totMedals$tot_medals)

#sathish ----

Reg_Pop<-lm(formula = tot_medals~population,data=Pop_totMedals)
summary(Reg_Pop)

#  Assuming an alpha of 0.05, the p-value of 0^-6 indicates that population does have a correlation to total medals won.Of note is that the p-value for BMI & population is smaller than age, so we will run the regression for age, BMI & population as out independent variables and analyze the output.
```


```{r}
#multilinear Regression to predict toal medals from BMI, Age and Population

#sathish -----
pairs(~tot_medals+bmi_metric+age+population, data = athletes)  # scotter plot
#sathish -----

Reg_Combined<-lm(formula = tot_medals~bmi_metric+age+population,data=athletes)
summary(Reg_Combined)

# From the regression output, population and BMI have a p-value below our alpha of 0.05 indicating they have a stronger correlation to total medals won than age.

```

```{r}
#Multi-linear regression to predict total medals from the average age and sport



All_Sports<- athletes %>%
  group_by(sport) %>%
  summarise(age = mean(age), tot_medals = sum(tot_medals))

All_Sports$rank<- c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28) #add ranking order

#sathish -----
pairs(~tot_medals+age+rank, data = All_Sports)  # scotter plot
#sathish -----


agesport_Reg<-lm(formula = tot_medals~age+rank,data=All_Sports)
summary(agesport_Reg)

#
```










```{r}

countrieswithmedals <- athletes %>% filter(tot_medals>0) %>%  group_by(country) %>% summarize(medals=sum(tot_medals,na.rm = TRUE), gold = sum(gold),silver = sum(silver),bronze = sum(bronze))
# countrieswithmedals

countriesbyRank <- countrieswithmedals[order(-countrieswithmedals$medals),]
countriesbyRank <- countriesbyRank %>% mutate(rank = 1:n())
col_order <- c("rank","country","medals","gold","silver","bronze")
# countriesbyRank
countriesbyRank <- countriesbyRank[, col_order]

Top5 <- data.frame(countriesbyRank[1:5,],stringsAsFactors = FALSE )
Top5$country <- recode(Top5$country
                                ,'United States' = 'USA'
                                ,'United Kingdom' = 'UK'
                                )
# Top5

map.world <- map_data("world")

glimpse(countriesbyRank)
as.factor(Top5$country) %>% levels()


df.country_points <- as.data.frame(Top5$country,stringsAsFactors = F)
colnames(df.country_points) <- c("country")
df.country_points$country <- as.character(df.country_points$country)

# ?register_google

register_google("AIzaSyDWp5xj2USZ5eOYnL7V5sJMx0w76d1NCDY")
    
df.country_points <- mutate_geocode(df.country_points, country)
# glimpse(df.country_points)

df.country_points <- left_join(df.country_points,Top5)
# df.country_points


map.world_olympics <- left_join(map.world, Top5, by = c('region' = 'country')) %>% mutate(fill_flg = ifelse(is.na(rank),F,T))
map.world_olympics <- rename(map.world_olympics, c("country"="region"))

map.world_olympics <- map.world_olympics %>% mutate(fill_flg = ifelse(fill_flg=="TRUE", country, fill_flg))

map.world_olympics$fill_flg <- recode(map.world_olympics$fill_flg,'FALSE' = 'Others')

#rename column name
names(map.world_olympics)[names(map.world_olympics)=="fill_flg"] <- "Countries"

map.world_olympics
unique(map.world_olympics[c("Countries")])
# df.country_points
```


```{r}

#map with dotted plot
top5.map <- ggplot() + geom_polygon(data = map.world_olympics, aes(x = long, y = lat, group = group),show.legend = FALSE,fill="white",color="#CCCCCC") 
top5.map <- top5.map +  geom_point(data = df.country_points, aes(x = lon, y = lat,color=country,size=order(-rank)) )
        
top5.map +  labs(x = NULL, 
                  y = NULL, 
                  title = " Top 5 Countries by Number of medals", 
                  caption = "Data: Olympics, 2016") + theme () 


#Colour Palette

pal <- c(
  "USA" = "red",
  "Germany" = "orange", 
  "UK" = "yellow", 
  "Russia" = "forestgreen" ,
  "China" ="blue",
  "Others" ="#CCCCCC"
)

mapG <- ggplot() + geom_polygon(data = map.world_olympics, aes(x = long, y = lat, group = group,fill=Countries),show.legend = TRUE)  + scale_fill_manual( values = pal, limits = names(pal) )
mapG <- mapG +  
                 labs(x = NULL, 
                  y = NULL, 
                  title = " Top 5 Countries by Number of medals", 
                  caption = "Data: Olympics, 2016") +theme 

mapG


```





