---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

-----------------------------------------------------------------------------------
Title: "IST687 – Writing functions and doing some initial data analysis"
Name: Sathish Kumar Rajendiran
Week: 2
Date: 04/18/2020
-----------------------------------------------------------------------------------
Exercise: 
      Explore the mtcars dataset (which is already included in R). Copy the mtcars dataset into a new variable (called it myCars), so that if you mess up, you can start again very easily (by copying mtcars into myCars again)
      
      { }
Step 0: Clone the dataset into new Dataframe.

```{r}
# clone mtcars into mycarsdf dataframe
  mycarsdf <- data.frame(mtcars)
  head(mycarsdf,10)
  
  mycarsdf <- cbind(name = rownames(mycarsdf), mycarsdf)
  rownames(mycarsdf) <- 1:nrow(mycarsdf)
  head(mycarsdf,10)
```
  
Step 1: What is the hp (hp stands for “horse power”)

```{r}

#   1) What is the highest hp?
      maxHp <- max(mycarsdf$hp)
      sprintf("%s is the highest horse power!", maxHp)
      
#   2) Which car has the highest hp?
      maxHp_index <- which.max(mycarsdf$hp)
      print(mycarsdf[maxHp_index,1],max.levels = 0)
      #Maserati Bora
      
```

Step 2: Explore mpg (mpg stands for “miles per gallon”)

```{r}

#   3) What is the highest mpg?

      maxMPG <- max(mycarsdf$mpg)
      sprintf("%s is the highest mpg!", maxMPG)

#   4) Which car has the highest mpg?
      maxMPG_index <- which.max(mycarsdf$mpg)
      print(mycarsdf[maxMPG_index,1],max.levels = 0)     
      #Toyota Corolla
  
#   5) Create a sorted dataframe, based on mpg    
      
      carsDFbyMPG <- mycarsdf[order(-mycarsdf$mpg),]
      head(carsDFbyMPG,5)
      tail(carsDFbyMPG,5)
      
```

Step 3: Which car has the “best” combination of mpg and hp?
```{r}
#  6) What logic did you use?

# lets divide hp by mpg to find the best value car
    mycarsdf$hpBympg <- mycarsdf$hp/mycarsdf$mpg

# re-arrange columns
    bestcarsDF <- mycarsdf[c("name","mpg","hp","hpBympg","cyl","disp","drat","wt","qsec","vs","am","gear","carb")]
    bestcarsDF
    
#  7) Which car?
# find the car that has best of value
    bestcar_index <- which.max(bestcarsDF$hpBympg)
    print(bestcarsDF[bestcar_index,1],max.levels = 0) 
    
```

Step 4: Which car has “best” car combination of mpg and hp, where mpg and hp must be given equal weight?
```{r}
    
    bestcarsDF$scalempg <- scale(bestcarsDF$mpg)
    bestcarsDF$scalehp  <- scale(bestcarsDF$hp)
    cols <- c('scalempg','scalehp')
    bestcarsDF$mostvalued <- apply(bestcarsDF[,cols],1,sum)
    
    finaldf <- bestcarsDF[c("name","mpg","hp","hpBympg","scalempg","scalehp","mostvalued")]
    finaldf <- finaldf[order(-finaldf$mostvalued),]
    
    mostvalued_index <- which.max(finaldf$mostvalued)
    print(finaldf[mostvalued_index,1],max.levels = 0) 
    #Maserati Bora
```


